---
title: "HW_1- DATA VISUALIZATION"
by: Jack Wright
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:


## Rank

This is the rank of the company, presumably some combination of revenue and growth rate. A visual analysis of this will be best to decipher how the ranking is done. 

## Growth Rate

This is the rate at which the companies are growing. Let us note that smaller companies will be more sensitive to change in size.

## Revenue

This will be how much money the company is pulling in. keep in mind this isn't profit. We will see if this is relevant later.

## City/State

There might be some correlation between location, we should be careful about developing causal relationships with location because it might be due to factors outside this dataset
```{r}
library(tidyverse)
#number of industries
inc$Industry<-as.factor(inc$Industry)
inc%>%count(Industry, sort=TRUE)
```
My first instinct is to say that "newer" industries make up a higher percentage of fast growing companies because of `IT Services`, but note that Software is in the middle of the pack. It is also hard to argue that `Business Products & Services` and `Advertising` are "new". 


Maybe you could say that companies that provide services as opposed to products top the list. 

## summary by industry


Note that the percent of the list taken up by an industry does not translate to revenue. Keep in mind that revenue does not equal profit. A company providing a service might make more profit because they don't have to account for the cost of a product, and they might have less employees to pay for in order to produce the product. These things will be more easily investigated with graphs.
```{r}
inc%>%group_by(Industry)%>%summarize(growth_rate=round(mean(Growth_Rate),1),revenue=mean(Revenue), employees=round(mean(Employees),1))%>%arrange(desc(revenue))
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}

library(ggplot2)
library(ggthemes)
#convert State variable to factor
inc$State<-as.factor(inc$State)
df_state<-inc%>%count(State, sort=TRUE)
ggplot(df_state, aes(x=n, y=reorder(State,n)))+geom_bar(stat='identity')+xlab('businesses')+ylab('state')+theme_tufte(base_size = 11, base_family = "serif", ticks = TRUE)+theme(aspect.ratio = 1)+ggtitle('Count of Businesses by state in "fastest 5000"')
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# Answer Question 2 here

#find 3rd highest state

state_3<-as.character(df_state$State)[3]
df_state_3<-inc%>%filter(State==state_3)%>%na.omit()

ggplot(df_state_3, aes(x=log(Employees),y=reorder(factor(Industry),Employees, FUN=median)))+geom_boxplot()+theme_tufte(base_size = 11, base_family = "serif", ticks = TRUE)+labs(x='log(Employees)',y='Industry',subtitle =paste('plot of Employees by industry in',state_3))
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here

#create revenue per employee statistic

df_rev_emp<-inc%>%drop_na(c(Employees,Revenue))%>%select(Industry,Employees,Revenue)%>%group_by(Industry)%>%summarize(revenue=sum(Revenue),employees=sum(Employees))

df_rev_emp<-inc%>%
  drop_na(c(Employees,Revenue))%>%
  select(Industry,Employees,Revenue)%>%
  group_by(Industry)%>%
  summarize(revenue=sum(Revenue),employees=sum(Employees))

df_rev_emp<-inc%>%
  drop_na(c(Employees,Revenue))%>%
  select(Industry,Employees,Revenue)%>%
  mutate(rev_emp=Revenue/Employees)

ggplot(df_state_3, aes(x=log(Employees),y=reorder(factor(Industry),Employees, FUN=median)))+geom_boxplot()+theme_tufte(base_size = 11, base_family = "serif", ticks = TRUE)+labs(x='log(Employees)',y='Industry',subtitle =paste('plot of Employees by industry in',state_3))

ggplot(df_rev_emp, aes(x=log(rev_emp), y=reorder(factor(Industry),rev_emp, FUN=median)))+geom_boxplot()+theme_tufte(base_size = 11, base_family = "serif", ticks = TRUE)+labs(x='log(Revenue)',y='Industry', subtitle = 'Revenue per Employee by Industry')
```


